---
title: "Parallel Computing"
author: "Lihle Sigenu"
format: pdf
editor: visual
---

# Question 1

Use a `foreach` loop to repeat 100 times:

generate a random sample from an exponential distribution with mean 1

calculate mean and variance

row-bind your results (rbind) (results = mean and variance).

```{r}
library(foreach)

# define combine technique and add in "", rbind so each run is computed by row
rand_sample <- foreach(i = 1:100, .combine = 'rbind') %do% {
  sample <- rexp(10, 1)
  # bind the mean and var into one row
  sample_stats <- c(mean(sample), var(sample))
}
head(rand_sample)
```

# Question 2

Use the `doParallel` package and `foreach` to bootstrap the median for the galaxies data (in library MASS).

If the `foreach` function needs access to data or a function from a certain package, this can be done by adding the `.packages='MASS'` (for example) argument.

How does processing time compare to that of serial processing? If each iteration’s run time is small, relative to the amount of data that needs to be loaded and returned, parallel processing might not actually speed up the total run time. Bootstrapping is relatively small: draw a sample, calculate a statistic. It might only start making a difference if each chunk becomes large relatively to the overheads of data transfer. Experiment with this. Try doing 1000 bootstrap samples at a time instead of managing single bootstrap samples.

```{r}
library(doParallel)
library(MASS)
head(galaxies)

cl <- makeCluster(3)
registerDoParallel(cl)
gals <- galaxies

#Sequential time estimation
system.time({
b_medians <- foreach(i = 1:1000, .combine = "rbind", .packages='MASS') %do% {
 b_sample <-  sample(galaxies, replace= T)
 median <- c(median(b_sample))
}
})

# Parallel time estimation
system.time({
b_medians <- foreach(i = 1:1000, .combine = "rbind", .packages='MASS') %dopar% {
 b_sample <-  sample(galaxies, replace= T)
 median <- c(median(b_sample))
}
})
stopCluster(cl)
```

# Question 3

Estimate coverage of a percentile bootstrap confidence interval for the following scenario: sample of size 50 from an exponential distribution with mean 1.

Coverage is basically the number of confidence intervals that contain the true parameter value, here the true mean of the exponential is 1. Want to compute

```{r}
library(foreach)
library(doParallel)
reps <- 1000 
B <- 1000
alpha <- 0.05
n <- 50
mean <- 1

c1 <- makeCluster(2)
registerDoParallel(c1)

coverage <- foreach(i = 1:reps, .combine = c) %dopar% {
  # obtain original sample from exponential dist
  original <- rexp(50,1)
  
  # bootstrap means using replicate B times
  b_means <- replicate(B, mean(sample(original, size = n, replace = T)))
  
  # confint
  confint <- quantile(b_means, c(alpha/2, 1-alpha/2))
  
  # determine whether the confidence interval contains the true mean = 1
  # if lower bound <= mean AND mean <= upper bound
  as.numeric(confint[1] <= mean & mean <= confint[2])
}

cat("The mean coverage is: ", mean(coverage))
stopCluster(c1)
```

# Question 4

The package iterators provides several functions that can be used to create sequences for the `foreach` function. For example, the `irnorm` function creates an object that iterates over vectors of normally distributed random numbers. It is useful when you need to use random variables drawn from one distribution in an expression that is run in parallel.

In this exercise, use the `foreach` and `irnorm` functions to iterate over 3 vectors, each containing 5 random variables. Find the largest value in each vector, and print those largest values.

Before running the `foreach` function set the seed to 1234.

```{r}
set.seed(1234)
library(iterators)

c3 <- makeCluster(3)
registerDoParallel(c3)


largest <- foreach(norms = irnorm(5, count = 3), .combine = c) %dopar% {
  max(norms)
}

print(largest)

```

# Question 5

Compare run time between `parLapply`, `foreach` and `replicate` for the above problem.

```{r}
library(iterators)

c3 <- makeCluster(3)
registerDoParallel(c3)

###########
# parLapply
###########

set.seed(1234)

# norms list stores the 3 vectors in a list, lapply does the function 3 times, function(i) is to repeat 1:3 times, rnorm generates 5 random numbers
norms_list <- lapply(1:3, function(i) rnorm(5))
# parLapply computes the max function to the list generated above, using the clients in cl
parLapply_time <- system.time({
  largest_parLapply <- parLapply(c3, norms_list, max)
})
cat("The parLapply time is: ", "\n")
print(parLapply_time, "\n", " ", "\n")

########
#foreach
########

set.seed(1234)

foreach_time <- system.time({
largest_foreach <- foreach(norms = irnorm(5, count = 3), .combine = c) %dopar% {
  max(norms)
}
})
cat("The foreach time is: ", "\n")
print(foreach_time, "\n")

##########
#replicate
##########

set.seed(1234)

replicate_time <- system.time({
  largest_replicate <- replicate(3, max(rnorm(5)))
})
cat("The replicate time is: ", "\n")
print(replicate_time, "\n")

stopCluster(c3)
```
